---
title: "Untitled"
author: "Brooks Coetzee III"
date: "2022-09-21"
output: html_document
---

```{r}
### Packages

library(httr) # httr is organised around the six most common http verbs: GET(), PATCH(), POST(), HEAD(), PUT(), and DELETE().
library(rvest) # Easily Harvest (Scrape) Web Pages
library(tidyverse)
library(jsonlite)  ## These functions are used to convert between JSON data and R objects.
library(lubridate)  ## Lubridate provides tools that make it easier to parse and manipulate dates.
library(readxl)
library(ggplot2)
library(xgboost)
library(randomForest)
```
Definition of variables
*FPTS: Total Fantansy for that season
*FPPG: Fantasy Points per game
*GP: Games Played
*TPA: Total Plate Appearances
*AB: At bats
*R: Runs
*H: Hits
*1B,2B,3B: Singles, Doubles, Triples
*HR: Homeruns
*RBI: Runs Batted In
*AVG: Batting Average
*OBP: On Base Percentage
*SLG: Slugging
*OPS: One base plus slugging
*BB: Walks
*SO: Strike Outs
*TB: Total Bases
*SB: Stolen Bases
*CS: Caught Stealing
*HBP: Hit By Pitch
*SH: Sac Hit
*SF: Sac Fly
*E: Error
*A: Assists
*GDP: Grounded into Double Play


2021 Season Stats
```{r}
baseball_2021 <- read_excel('Baseball_Data_2021.xlsx')
summary(baseball_2021)
head(baseball_2021)
tail(baseball_2021)

sum(is.na(baseball_2021))
#0 NA values
```
Density Plots:
```{r}
# Create plot
g_1 <- ggplot(baseball_2021, aes(x = FPTS)) + # Set X-axis as insurance charges
  geom_density(fill = "blue", alpha = 0.3) + # Use geom_density to get density plot
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(x = "FPTS", # Set plot labels
       title = "Density of FPTS")

g_1 # Generate plot

#Great density in the 400 for PTS on the year
```
relation between FPPG and AVG
```{r}
g_3 <- ggplot(baseball_2021, # Set dataset 
              aes(y = FPPG, # Set y-axis as insurance charges 
                  x = AVG)) + # Set x-axis as age.
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(y = "FPPG", # Set plot labels
       x = "AVG",
       title = "FPPG and AVG")

g_3

#There is a positive correlation with a wide standard error. this shows that there is a correlation between AVG and fantasy points per game.
```


2022 Season
```{r}
#as of September 20, 2022
baseball_2022 <- read_excel('Baseball_Data_2022_Updated.xlsx')
summary(baseball_2022)
head(baseball_2022)
tail(baseball_2022)
sum(is.na(baseball_2022))
#0 NA values
```
Density Plot:
```{r}
g_2 <- ggplot(baseball_2022, aes(x = FPTS)) + # Set X-axis as insurance charges
  geom_density(fill = "blue", alpha = 0.3) + # Use geom_density to get density plot
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(x = "FPTS", # Set plot labels
       title = "Density of FPTS")

g_2 # Generate plot

#Great density in the 400 for PTS on the year
#Lower scoreres overall and lower number of players over 600 points compared to the previous year.
```
relation between FPPG and AVG:
```{r}
g_4 <- ggplot(baseball_2022, # Set dataset 
              aes(y = FPPG, # Set y-axis as insurance charges 
                  x = AVG)) + # Set x-axis as age.
  geom_point(color = "blue", alpha = 0.3) + # Use geom_point to get scatter plot
  geom_smooth(method = "lm") + # Add smoothing line
  theme_bw() + # Set theme for plot
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  labs(y = "FPPG", # Set plot labels
       x = "AVG",
       title = "FPPG and AVG")

g_4

#There is a positive correlation with a wide standard error. this shows that there is a correlation between AVG and fantasy points per game.
# Like the previous graph there is a lower correlation between the two variables than the previous year.
```
To conclude from our initial findings, The most broad offensive statistic has a correlation to fantasy points. However it is a slight positive relation. Our goal of this project is to see if we can predict what stats causes the player to set them apart from the rest of the players. 
--------------------------------------------------------------------------------

# Lag points

```{r}
head(baseball_2021)
head(baseball_2022)

points_2022 <- rep(NA, nrow(baseball_2021))
for(i  in 1:nrow(baseball_2021)){
  if(baseball_2021$PLAYER[i] %in% baseball_2022$PLAYER){
    points_2022[i] <- baseball_2022$FPTS[which(baseball_2022$PLAYER == baseball_2021$PLAYER[i])]
  }
}

points_2022

baseball_2021$point_2022 <- points_2022

baseball_lagged <- na.omit(baseball_2021)

baseball_lagged[baseball_lagged == "—"] <- "0"
baseball_lagged <- as.data.frame(baseball_lagged)
for(i in 2:ncol(baseball_lagged)){
  baseball_lagged[,i] <- as.numeric(baseball_lagged[,i])
}

summary(baseball_lagged)
```
53 players present in both data sets




Analysis of 2022 to 2021
# Initial Regression model:
```{r}
#Total Fantasy Points with Significant Predictors
noplayer_2021 <- baseball_lagged[,-c(1)]
model_2021 <- lm(point_2022 ~ SO + AVG + OPS + SLG + R + H + TB + GP + HR + FPTS ,data = noplayer_2021)
summary(model_2021)

all <- lm(point_2022 ~ ., data = noplayer_2021)

summary(all)

```
# Visuals

XGboost
```{r}
set.seed(111111)
test_index <- sample(1:nrow(baseball_lagged), nrow(baseball_lagged) * 0.2)

dtrain <- xgb.DMatrix(data = as.matrix(baseball_lagged[-test_index,c("SO", "AVG", "OPS", "SLG",
                                                                      "R", "H",  "TB", "GP", "HR",
                                                                      "FPTS")]), label = as.numeric(baseball_lagged$point_2022[-test_index]))
# Create test matrix
dtest <- xgb.DMatrix(data = as.matrix(baseball_lagged[test_index,c("SO", "AVG", "OPS", "SLG",
                                                                      "R", "H",  "TB", "GP", "HR",
                                                                      "FPTS")]), label = as.numeric(baseball_lagged$point_2022[test_index]))


bst_1 <- xgboost(data = dtrain, # Set training data

               

               nrounds = 100, # Set number of rounds

               

               verbose = 1, # 1 - Prints out fit

                print_every_n = 20# Prints out result every 20th iteration

 )

```
```{r}
# Extract importance
imp_mat <- xgb.importance(model = bst_1)
# Plot importance (top 10 variables)
xgb.plot.importance(imp_mat, top_n = 10)
```


